Analysis: The Sound of Teaching Music 1

### 2019-10-04 script/RNotebook/articulation.Rmd
add order analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@

- [Summary](#art_summary)

@@

@@

- [Summary](#dyn_summary)

@@

# Calculate KOR for each interval
df_kor <- data.frame()
for (subnr in unique(df_kot$SubNr)){
  for (block in unique(df_kot$BlockNr[df_kot$SubNr == subnr])){
    for (trial in unique(df_kot$TrialNr[df_kot$SubNr == subnr & df_kot$BlockNr == block])){
      df_current <- df_kot %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
      df_current$KOR <- df_current$KOT/ioi$Mean[ioi$SubNr == subnr & ioi$BlockNr == block & ioi$TrialNr == trial]
      df_kor <- rbind(df_kor, df_current)
    }
  }
}

@@

```{r kot_dyn_bar,  echo = FALSE}
p_kot_dyn <- ggplot(data = kot_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_dyn
```

@@

>>>>

- Order analysis (teaching/performing first?)
  + [All](#kot_art_order_group_all)

@@

- Order analysis (teaching/performing first?)
  + [All](#kor_art_order_group_all)

@@

- Order analysis (teaching/performing first?)
  + [All](#kot_dyn_order_group_all)

@@

- Order analysis (teaching/performing first?)
  + [All](#kor_dyn_order_group_all)

@@

# For each individual, each block, each trial
ioi <- aggregate(IOI~SubNr*BlockNr*TrialNr*Condition*Skill, data = df_ioi,
                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
ioi <- cbind(ioi[,1:5], as.data.frame(ioi[,6]))
# Change colnames
colnames(ioi) <- c("SubNr", "BlockNr", "TrialNr", "Condition", "Skill", "N", "Mean", "SD")

# Calculate KOR for each interval

df_kor <- data.frame()
for (subnr in unique(df_kot$SubNr)){
  for (block in unique(df_kot$BlockNr[df_kot$SubNr == subnr])){
    for (trial in unique(df_kot$TrialNr[df_kot$SubNr == subnr & df_kot$BlockNr == block])){
      df_current <- df_kot %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
      df_current$KOR <- df_current$KOT/ioi$Mean[ioi$SubNr == subnr & ioi$BlockNr == block & ioi$TrialNr == trial]
      df_kor <- rbind(df_kor, df_current)
    }
  }
}

@@

```{r kot_dyn_bar,  echo = FALSE}
p_kot_dyn <- ggplot(data = kot_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn
```

@@

## KOT-Order-Articulation
- Investigate whether there is any order effect of the conditions.

### 1. All{#kot_art_order_group_all}
```{r kot_art_order_all, echo = FALSE}
# For each participant
kot_art_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
kot_art_order <- cbind(kot_art_order[,1:5], as.data.frame(kot_art_order[,6]))
# Change colnames
colnames(kot_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")

# Group mean
kot_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_art_order,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kot_art_order_all <- cbind(kot_art_order_all[,1:4], as.data.frame(kot_art_order_all[,5]))
# Change colnames
colnames(kot_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
print(kot_art_order_all)
```

### Plot
#### Bar graph
- Error bars show SEM of the mean

```{r kot_art_order_all_bar, echo = FALSE}
p_kot_art_order_all <- ggplot(data = kot_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  facet_wrap(Order ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_order_all
```

##### Box plot
```{r kot_art_order_all_box, echo = FALSE}
p_kot_art_order_all_box <- ggplot(data = kot_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
  facet_wrap(Order ~ .) +
  labs(y = "KOT", subtitle = "KOT: Articulation")
p_kot_art_order_all_box
```

#### Statistics
##### Normality check
```{r normality5, echo = FALSE}
# Normality check
kot_art_order$Subcomponent <- factor(kot_art_order$Subcomponent)
kot_art_order_norm <- by(kot_art_order$Mean, list(kot_art_order$Condition, kot_art_order$Subcomponent, kot_art_order$Order), shapiro.test)
print(kot_art_order_norm)
```

##### Three-way ANOVA (ez)
```{r stats9, echo = FALSE}
kot_art_order_aov <- ezANOVA(
  data = df_kot_art
  , dv = .(KOT)
  , wid = .(SubNr)
  , between = .(Order)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_art_order_aov$ANOVA$F <- round(kot_art_order_aov$ANOVA$F, 4)
kot_art_order_aov$ANOVA$p <- round(kot_art_order_aov$ANOVA$p, 4)
print(kot_art_order_aov)
```

## KOR-Order-Articulation
- Investigate whether there is any order effect of the conditions.

### 1. All{#kor_art_order_group_all}
```{r kor_art_order_all, echo = FALSE}
# For each participant
kor_art_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
kor_art_order <- cbind(kor_art_order[,1:5], as.data.frame(kor_art_order[,6]))
# Change colnames
colnames(kor_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")

# Group mean
kor_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_art_order,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kor_art_order_all <- cbind(kor_art_order_all[,1:4], as.data.frame(kor_art_order_all[,5]))
# Change colnames
colnames(kor_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
print(kor_art_order_all)
```

### Plot
#### Bar graph
- Error bars show SEM of the mean

```{r kor_art_order_all_bar, echo = FALSE}
p_kor_art_order_all <- ggplot(data = kor_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  facet_wrap(Order ~ .) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_order_all
```

##### Box plot
```{r kor_art_order_all_box, echo = FALSE}
p_kor_art_order_all_box <- ggplot(data = kor_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
  facet_wrap(Order ~ .) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_order_all_box
```

#### Statistics
##### Normality check
```{r normality6, echo = FALSE}
# Normality check
kor_art_order$Subcomponent <- factor(kor_art_order$Subcomponent)
kor_art_order_norm <- by(kor_art_order$Mean, list(kor_art_order$Condition, kor_art_order$Subcomponent, kor_art_order$Order), shapiro.test)
print(kor_art_order_norm)
```

##### Three-way ANOVA (ez)
```{r stats10, echo = FALSE}
kor_art_order_aov <- ezANOVA(
  data = df_kor_art
  , dv = .(KOR)
  , wid = .(SubNr)
  , between = .(Order)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_art_order_aov$ANOVA$F <- round(kor_art_order_aov$ANOVA$F, 4)
kor_art_order_aov$ANOVA$p <- round(kor_art_order_aov$ANOVA$p, 4)
print(kor_art_order_aov)
```

## KOT-Order-Dynamics
- Investigate whether there is any order effect of the conditions.

### 1. All{#kot_dyn_order_group_all}
```{r kot_dyn_order_all, echo = FALSE}
# For each participant
kot_dyn_order <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
kot_dyn_order <- cbind(kot_dyn_order[,1:5], as.data.frame(kot_dyn_order[,6]))
# Change colnames
colnames(kot_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")

# Group mean
kot_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kot_dyn_order,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kot_dyn_order_all <- cbind(kot_dyn_order_all[,1:4], as.data.frame(kot_dyn_order_all[,5]))
# Change colnames
colnames(kot_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
print(kot_dyn_order_all)
```

### Plot
#### Bar graph
- Error bars show SEM of the mean

```{r kot_dyn_order_all_bar, echo = FALSE}
p_kot_dyn_order_all <- ggplot(data = kot_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  facet_wrap(Order ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn_order_all
```

##### Box plot
```{r kot_dyn_order_all_box, echo = FALSE}
p_kot_dyn_order_all_box <- ggplot(data = kot_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
  facet_wrap(Order ~ .) +
  labs(y = "KOY", subtitle = "KOT: Dynamics")
p_kot_dyn_order_all_box
```

#### Statistics
##### Normality check
```{r normality7, echo = FALSE}
# Normality check
kot_dyn_order$Subcomponent <- factor(kot_dyn_order$Subcomponent)
kot_dyn_order_norm <- by(kot_dyn_order$Mean, list(kot_dyn_order$Condition, kot_dyn_order$Subcomponent, kot_dyn_order$Order), shapiro.test)
print(kot_dyn_order_norm)
```

##### Three-way ANOVA (ez)
```{r stats11, echo = FALSE}
kot_dyn_order_aov <- ezANOVA(
  data = df_kot_dyn
  , dv = .(KOT)
  , wid = .(SubNr)
  , between = .(Order)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_dyn_order_aov$ANOVA$F <- round(kot_dyn_order_aov$ANOVA$F, 4)
kot_dyn_order_aov$ANOVA$p <- round(kot_dyn_order_aov$ANOVA$p, 4)
print(kot_dyn_order_aov)
```

## KOR-Order-Dynamics
- Investigate whether there is any order effect of the conditions.

### 1. All{#kor_dyn_order_group_all}
```{r kor_dyn_order_all, echo = FALSE}
# For each participant
kor_dyn_order <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57), FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x)), 4)})
kor_dyn_order <- cbind(kor_dyn_order[,1:5], as.data.frame(kor_dyn_order[,6]))
# Change colnames
colnames(kor_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD")

# Group mean
kor_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = kor_dyn_order,
                                     FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
kor_dyn_order_all <- cbind(kor_dyn_order_all[,1:4], as.data.frame(kor_dyn_order_all[,5]))
# Change colnames
colnames(kor_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM")
print(kor_dyn_order_all)
```

### Plot
#### Bar graph
- Error bars show SEM of the mean

```{r kor_dyn_order_all_bar, echo = FALSE}
p_kor_dyn_order_all <- ggplot(data = kor_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  facet_wrap(Order ~ .) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_order_all
```

##### Box plot
```{r kor_dyn_order_all_box, echo = FALSE}
p_kor_dyn_order_all_box <- ggplot(data = kor_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) +
  facet_wrap(Order ~ .) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_order_all_box
```

#### Statistics
##### Normality check
```{r normality8, echo = FALSE}
# Normality check
kor_dyn_order$Subcomponent <- factor(kor_dyn_order$Subcomponent)
kor_dyn_order_norm <- by(kor_dyn_order$Mean, list(kor_dyn_order$Condition, kor_dyn_order$Subcomponent, kor_dyn_order$Order), shapiro.test)
print(kor_dyn_order_norm)
```

##### Three-way ANOVA (ez)
```{r stats12, echo = FALSE}
kor_dyn_order_aov <- ezANOVA(
  data = df_kor_dyn
  , dv = .(KOR)
  , wid = .(SubNr)
  , between = .(Order)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_dyn_order_aov$ANOVA$F <- round(kor_dyn_order_aov$ANOVA$F, 4)
kor_dyn_order_aov$ANOVA$p <- round(kor_dyn_order_aov$ANOVA$p, 4)
print(kor_dyn_order_aov)
```
